---
title: "Conclusion"
format:
  html: 
    toc: false
    link-external-icon: false
    link-external-newwindow: false
---

```{r include = FALSE}
source(here::here("R", "buttons.R"))
```

Now that you've completed this module, you should have a good handle on the fundamentals of pipes in R. Specifically,
you should have a good sense of the following concepts.

- What problems pipes solve and which situations benefit from a pipeline-based approach.
- How you can use pipes to compose multiple functions into a single pipeline.
- How and when to use four different flavors of pipe operator.

## Next Steps

Pipes are very useful tools in many situations, but they truly shine when using in data processing workflows, especially
when combined with **dplyr** data manipulation functions. As such, we recommend the *Data Manipulation* module as a
natural next step that directly extends the tools you learned in this module.

```{r}
#| echo: false
#| results: asis
# next_module_button("https://www.openstatprog.com/data-manipulation/")
next_module_button("/in_progress.qmd")
```
