---
title: "Usage"
webr:
  packages: 'dplyr'
---

```{webr-r}
#| autorun: true
#| context: setup

dir.create("data")
download.file("https://openstatprog.com/raw-data/bfi.rds", "data/bfi.rds")

bfi <- readRDS("data/bfi.rds")
```

```{r, include = FALSE}
bfi <- readRDS(here::here("data", "bfi.rds"))
```

## How Pipes Work

The pipe operator simply passes the output from the upstream function into the first argument of the downstream
function. Hence, these two expressions are equivalent:

```{r}
mean(bfi$age)
bfi$age |> mean()
```

::: practice

Create a pipeline to complete the following operations on the `mtcars` dataset.

1. Standardize each variable in the `mtcars` dataset.
1. Compute the mean of each variable in the standardized `mtcars` data.

Your solution should be a single pipeline that produces a length-one logical vector.

:::: {.panel-tabset}

## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.hint >}}

You can use the following Base R functions in your solution:

- `scale()`: Standardize variables
- `colMeans()`: Compute the mean of each column in a matrix or data frame.

## {{< var tab_title.solution >}}

```{r}
mtcars |> scale() |> colMeans()
```

::::

:::

The data piped into a function only fill the first argument. We're free to specify any additional inputs in the normal
way. Hence, the following to expressions are also equivalent.

```{r}
var(bfi$a1, na.rm = TRUE)
bfi$a1 |> var(na.rm = TRUE)
```

::: practice

Create a pipeline to complete the following operations in one expression.

1. Create a random sample of 1000 normally distributed values with:
   - Mean of 10
   - SD of 5
1. Compute the range of your sample.
    - I.e., the difference between the largest and smallest value.
1. Check if the range of your sample is at least 1.5 times the population variance.

Your solution should be a single pipeline that produces a length-one logical vector.

:::: {.panel-tabset}

## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.hint >}}

You probably want to use the following functions somewhere in your pipeline.

- `rnorm()`
- `range()`
- `diff()`

## {{< var tab_title.solution >}}

```{r}
m <- 10
s <- 5

rnorm(1000, mean = m, sd = s) |>
  range() |>
  diff() > (1.5 * s^2)
```

::::

:::

### What if the dataset isn’t the first argument?

Many R functions, particularly those that use a so-called *formula interface*, don’t take the input data as their first
argument. If you try to include such a function in a normal pipeline, it won’t work. For example, in the following code,
we try to use the `bfi` dataset to estimate a linear regression model wherein `age` and `open` predict `extra`.

```{r, error = TRUE}
bfi |> lm(extra ~ age + open)
```

In these cases, you can use the special placeholder token `_` to tell R explicitly where to insert the piped object.

```{r}
bfi |> lm(extra ~ age + open, data = _)
```

Note that we must name the argument to which we assign `_`. The following won't work.

```{r, error = TRUE}
bfi |> lm(extra ~ age + open, _)
```

This trick makes it possible to use nearly any function in a pipeline, even when the dataset isn't the first argument.


::: practice

Create a pipeline to perform the following operations with the `bfi` dataset:

1. Randomly sample $N = 500$ rows.
   - Use `dplyr::slice_sample()`.
1. Use the sampled rows to regress `agree` onto `extra`, `open`, and `gender`.
   - Use `lm()`.
1. Extract the residuals from the linear regression model.
1. Compute the sum of the absolute values of the residuals.

:::: {.panel-tabset}

## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.solution >}}

```{r}
library(dplyr)
set.seed(235711)

bfi |> 
  slice_sample(n = 500) |>
  lm(agree ~ extra + open + gender, data = _) |>
  resid() |>
  abs() |>
  sum()
```

*NOTES:*

- If you don't set the same (or any) seed, your code will produce a different value than the solution shown here
because your code will sample a different set of rows.
- If you set the same seed by running `set.seed(235711)` before your pipeline, you should get exactly the same result
because `slice_sample()` will use the same sequence of pseudo random numbers to pick the rows it "randomly" samples.

::::

:::
