---
title: "Motivation"
---

```{r, include = FALSE}
bfi <- readRDS(here::here("data", "bfi.rds"))
```

When you perform several transformations in R, your code can quickly become cluttered with nested parentheses or a
series of temporary variables. Pipes offer a cleaner, more intuitive alternative: they let you write code that mirrors
the logical flow of your analysis. Each step passes its output directly into the next, creating a readable sequence of
operations that tells the story of your data transformation from start to finish.

For example, consider using this sequence of operations to compute a vector of standard deviations with the `bfi` dataset:

1. Select five personality scale scores.
1. Compute their covariance matrix.
1. Extract the variances from this matrix.
1. Take the square roots to obtain standard deviations.

We could implement these calculations as the following nested function calls.

```{r}
sd <- sqrt(
  diag(
    cov(
      bfi[c("agree", "consc", "extra", "neuro", "open")],
      use = "pairwise"
    )
  )
)
sd
```

This approach does the job, but it's certainly not ideal (unless you happen to be a big fan of
[lisp](https://xkcd.com/297)). The main issue is readability: the code is difficult to parse because you need to read
the nested function calls from the inside out to understand the sequence of operations. The situation gets even worse if
we're not as careful about formatting our code.

```{r}
sd <- sqrt(diag(cov(bfi[c("agree", "consc", "extra", "neuro", "open")],
  use = "pairwise")))
sd
```

To make things clearer, we could break the analysis into several steps and save the intermediate results as a temporary
object that we repeatedly overwrite.

```{r}
tmp <- bfi[c("agree", "consc", "extra", "neuro", "open")]
tmp <- cov(tmp, use = "pairwise")
tmp <- diag(tmp)
sd <- sqrt(tmp)
sd
```

This approach is clearer than using deeply nested functions, but we still have notable problems. The repeated
assignments increase the opportunities for typos and bugs. For example, consider a situation where we ran the following
code earlier in our analysis.

```{r}
tmp <- bfi[ , 1:5]
```

Then we tried to execute our focal calculation using the temporary object approach, but we made a small typo.

```{r}
Tmp <- bfi[c("agree", "consc", "extra", "neuro", "open")]
tmp <- cov(tmp, use = "pairwise")
tmp <- diag(tmp)
sd2 <- sqrt(tmp)
sd2
```

Now, our vector contains the standard deviations of the five raw agreeableness items, not the intended scale scores.
However, our results is still a length-five numeric vector, so we could easily miss our mistake and use the wrong vector
in subsequent operations. For example, if we were intending to use these standard deviations to standardize the
personality scores, R would be happy to use either version.

```{r}
# Standardize using the correct SDs:
zDat1 <- scale(bfi[c("agree", "consc", "extra", "neuro", "open")], scale = sd)

# Standardize using the wrong SDs:
zDat2 <- scale(bfi[c("agree", "consc", "extra", "neuro", "open")], scale = sd2)

# Obviously, we get very different results:
head(zDat1 - zDat2, 10)
```

This type of error can be difficult to catch in real-world projects, and R won't do anything to warn you. So, we want to
minimize our opportunities for such mistakes, and pipes are a great way to do just that. Using a pipe, we can express
the same sequence of calculations in a single expression with a clear, logical structure.

```{r}
library(dplyr)

sd <- bfi |>
  select(agree, consc, extra, neuro, open) |>
  cov(use = "pairwise") |>
  diag() |>
  sqrt()
```

When we compose multiple function call using pipes, we call the resulting command a *pipeline*. Such pipelines solve
both the issues noted above.

1. It's easy to understand the sequence of calculations, because we can read off the analytic steps from left-to-right,
   top-to-bottom (just as you'd read text written in a European language, for example).
1. We do not need to make any temporary assignments, so we minimize our chances of typo-induced bugs.

In practice, pipes are most useful for implementing sequences of complex data processing steps. For example, suppose we
want to apply the following data processing steps to the `bfi` data.

1. Center the variable `age` on 18.
1. Create mean scores for `extraversion` and `neuroticism`.
1. Keep only participants aged 18 or older.
1. Select only the scale scores and demographic variables.
1. Sort the data by `extraversion` in ascending order.

Using pipes and `dplyr` functions, we can write:

```{r}
tmp1 <- bfi %>%
  mutate(age = age - 18,
         extra = rowMeans(across(matches("^e\\d$")), na.rm = TRUE),
         neuro = rowMeans(across(matches("^n\\d$")), na.rm = TRUE)) |>
  filter(age >= 0) |>
  select(extra, neuro, age, gender, education) |>
  arrange(extra)

head(tmp1, 20)
```

When we implement these computations as a pipeline, each step is explicit, readable, and self-contained: it’s as clear
as possible what happens and in what order. The same data processing written without pipes would be much harder to
follow:

```{r}
tmp2 <- bfi
tmp2$age   <- tmp2$age - 18
tmp2$extra <- rowMeans(tmp2[grep("^e\\d$", colnames(tmp2))], na.rm = TRUE)
tmp2$neuro <- rowMeans(tmp2[grep("^n\\d$", colnames(tmp2))], na.rm = TRUE)
tmp2 <- tmp2[tmp2$age >= 0, c("extra", "neuro", "age", "gender", "education")]
tmp2 <- tmp2[order(tmp2$extra), ]

head(tmp2, 20)
all.equal(tmp1, tmp2, check.attributes = FALSE)
```

Both approaches produce the same result, but the pipeline is far clearer.
