---
title: "Different Pipe Operators"
webr:
  packages: ['dplyr', 'magrittr']
---

```{webr-r}
#| autorun: true
#| context: setup

dir.create("data")
download.file("https://openstatprog.com/raw-data/bfi.rds", "data/bfi.rds")

bfi <- readRDS("data/bfi.rds")
```

```{r, include = FALSE}
bfi <- readRDS(here::here("data", "bfi.rds"))
```

## The **dplyr** Pipe: `%>%`

R's native pipe operator, `|>`, was only introduced in R 4.1. Before that, we used the analogous operator provided by
the **dplyr** package, `%>%`. The **dplyr** pipe is almost exactly the same as the base R version. For example, we can
re-create our first motivating example as follows.

```{r}
library(dplyr)

bfi[c("agree", "consc", "extra", "neuro", "open")] %>%
  cov(use = "pairwise") %>%
  diag() %>%
  sqrt()
```

For most practical purposes, the Base R and **dplyr** pipes are interchangeable, but there are a few minor differences
between `|>` and `%>%`. For example, the **dplyr** pipe uses the `.` character as the dataset placeholder symbol.

```{r}
mtcars %>%
  lm(mpg ~ hp + disp, data = .) %>%
  summary()
```

The **dplyr** pipe also allows more flexible usage of the dataset placeholder symbol.

```{r}
mtcars %>%
  lm(mpg ~ hp + disp, data = ., weights = .$wt) %>%
  summary()
```

```{r, error = TRUE}
mtcars |>
  lm(mpg ~ hp + disp, data = _, weights = _$wt) |>
  summary()
```

## The Exposition Pipe: `%$%`

The exposition pipe, `%$%`, is a very useful flavor or pipe operator provided by the **magrittr** package. The
exposition pipe offers an elegant alternative to the dataset placeholder symbol in many common types of pipeline. Most
notably, pipelines involving functions that employ the so-called *formula interface*, such as `lm()`, `t.test()`, or
`boxplot()`.

Instead of passing the processed dataset into the first argument of a downstream function, the exposition pipe *exposes*
the column names of the processed data frame to the downstream function. For example, the following code uses the
exposition pipe to regress `age` on `extra` in the `bfi` data.

```{r}
library(magrittr)

bfi %$% lm(extra ~ age + open)
```

The exposition pipe allows the `lm()` call to access the column names of `bfi` when resolving the model formula `extra ~
age`. We can achieve analogous results in several ways.

```{r}
# Explicitly regress vectors:
lm(bfi$extra ~ bfi$age + bfi$open)

# Use the Base R equivalent of %$%:
with(bfi, lm(extra ~ age + open))

# Use a dataset placeholder symbol:
bfi |> lm(extra ~ age + open, data = _)
```

::: practice

Use the `bfi` dataset and an exposition pipe to create a pipeline to draw a conditional boxplot that visualizes the
distribution of `agree` scores within `gender` groups.

- Use the Base R `boxplot()` function to draw the plot.

:::: {.panel-tabset}

## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.solution >}}

```{r}
library(magrittr)

bfi %$% boxplot(agree ~ gender)
```

::::

:::

::: practice

Create a pipeline that includes at least on exposition pipe to perform the following operations with the `bfi` dataset:

1. Randomly sample $N = 500$ rows.
   - Use `dplyr::slice_sample()`.
1. Use the sampled rows to regress `agree` onto `extra`, `open`, and `gender`.
   - Use `lm()`.
1. Extract the residuals from the linear regression model.
1. Compute the sum of the absolute values of the residuals.

:::: {.panel-tabset}

## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.solution >}}

```{r}
library(dplyr)
set.seed(235711)

bfi |> 
  slice_sample(n = 500) %$%
  lm(agree ~ extra + open + gender) |>
  resid() |>
  abs() |>
  sum()
```

::::

:::

## The Assignment Pipe: `%<>%`

The final type of pipe that we'll cover is certainly a niche tool, but it provides a nice little bit of syntactic sugar.
We only care about assignment pipes in particular types of pipelines. Names, pipelines with two characteristics:

1. The pipeline modifies a dataset
1. The pipeline overwrites the original dataset with the modified version

For example, the following pipeline uses **dplyr** data manipulation functions to modify the `bfi1` dataset and
overwrite the original with the modified version produced by the pipeline.

```{r}
bfi1 <- bfi
head(bfi1)

bfi1 <- bfi1 %>%
  slice_sample(n = 20) %>%
  select(-matches("\\d$"))

head(bfi1)
```

The assignment pipe cleans up the syntax a little but and makes the destructive nature of the pipeline clear. When you
see an assignment pipe, you know that the modified data produced by the pipeline will replace the original dataset.

```{r}
bfi1 <- bfi
head(bfi1)

bfi1 %<>%
  slice_sample(n = 20) %>%
  select(-matches("\\d$"))

head(bfi1)
```

Note that we must use the **dplyr** pipe, `%>%`, if we include an assignment pipe in our pipeline. The Base R pipe
doesn't play nicely with the assignment pipe. More specifically, the assignment pipe behaves like a basic pipe (i.e.,
`|>` or `%>%`) when combined with the Base R pipe operator, as you can see below.

```{r}
bfi1 <- bfi
head(bfi1)

bfi1 %<>%
  slice_sample(n = 20) |>
  select(-matches("\\d$"))

head(bfi1)
```

::: practice

Use the assignment pipe to create a pipeline that implements the following operations:

1. Standardize every column in the `mtcars` dataset.
1. Overwrites the original `mtcars` dataset with the standardized version.

:::: {.panel-tabset}

## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.hint >}}

You can use the `scale()` function to standardize variables.

## {{< var tab_title.solution >}}

```{r}
library(magrittr)

mtcars %<>% scale()
```

We can sanity check the results by computing the means and variances of each variable.

```{r}
colMeans(mtcars)
apply(mtcars, 2, var)
```

Looks good!

::::

:::
